---
title: "Final Project"
author: "Pranav Keshav and Hojin Ryoo"
date: "4-5-2020"
output: 
    html_notebook:
    fig_height: 6
    fig_width: 10
---

## Guiding Question

Are there any correlations between a pokemon's stats and type? Do these correlations make sense in the context of the type? 


## Data Access

Where: We found this dataset online. 

Who/When and Why: This dataset was created and is maintained by user eevee and some other contributors. This was collected in order to collect data from all different Pokemon games in order to see trends and make observations from the dataset. 

What: There are 964 cases in which the case is represented by the Pokemon name and then some data about it such as height, weight and base experience.

```{r}
library(DataComputing)
library(tidyverse)
```

```{r}
pk_names %>%
  summarise(total_cases = n())
```

```{r}
#At least two data sources are imported

pk_stats <- read_csv(file = "pokemon_stats.csv")
pk_types <- read_csv(file = "pokemon_types.csv")
pk_names_extra <- read_csv(file = "pokemon.csv")
stat_names <- read_csv(file = "stat_names.csv")
type_names <- read_csv(file = "type_names.csv")
pk_stats
pk_types
pk_names
stat_names
type_names
```

# Data Wrangling

```{r}
# Getting English type names
type_names <-
  type_names %>%
  filter(local_language_id == 9)
type_names
```

```{r}
# Getting English Stat Names
stat_names <-
  stat_names %>%
  filter(local_language_id == 9)
stat_names
```

```{r}
# Selecting the rows needed from the pokemon_names_extra var
pk_names <-
  pk_names_extra %>%
  select(species_id, identifier)
pk_names
```

```{r}
#Creating correct info

pk_stats_z <-
  pk_stats %>%
  left_join(y = pk_names, by = c("pokemon_id" = "species_id")) %>%
  select(identifier, stat_id, base_stat) %>%
  left_join(y = stat_names) %>%
  select(identifier, name, base_stat) %>%
  rename(stat = name, 
         name = identifier)
pk_stats_z
```

```{r}
#Creating id column
pk_stats_z$id = NA

pk_stats_z
```

```{r}
#Creating values for id column
for (row in 1:nrow(pk_stats_z)) 
{
  pk_stats_z$id[[row]] <- row
}

pk_stats_z
```

```{r}
for (row in 1:nrow(pk_stats_f))
{
  if (pk_stats_z$stat[[row]] == "Special Attack")
  {
    pk_stats_z$stat[[row]] <- "Special_Attack"
  }
  if (pk_stats_z$stat[[row]] == "Special Defense")
  {
    pk_stats_z$stat[[row]] <- "Special_Defense"
  }
}

pk_stats_z
```


```{r}
unclean <-
  pk_stats_z %>%
  spread(key = stat, value = base_stat)
pk_stats_z <-
  unclean %>%
  group_by(name) %>%
  summarise(HP = sum(HP, na.rm = TRUE), 
            Attack = sum(Attack, na.rm = TRUE),
            Defense = sum(Defense, na.rm = TRUE) ,
            Special_Attack = sum(Special_Attack, na.rm = TRUE),
            Special_Defense = sum(Special_Defense, na.rm = TRUE),
            Speed = sum(Speed, na.rm = TRUE))
pk_stats_z
```

```{r}
# Creating the master data
master2 <-
  pk_stats_z %>%
  left_join(pk_types_f) %>%
  select(name, type, slot, HP, Attack, Defense, Special_Attack, Special_Defense, Speed)
master2
```
