---
title: "Final Project"
author: "Pranav Keshav and Hojin Ryoo"
date: "4-5-2020"
output: 
    html_notebook:
    fig_height: 6
    fig_width: 10
---

## Guiding Question

Are there general trends in regards to pokemon stats and the correlation to what type they may be?


## Data Sources

We found this dataset online. 

This dataset was created and is maintained by user eevee and some other contributors. This was collected in order to collect data from all different Pokemon games in order to see trends and make observations from the dataset. 

There are 965 cases in which the case is represented by the Pokemon name and then some data about it such as height, weight and base experience.

We plan to use the base experience and relate it to the different types of Pokemon.


## Research

```{r}
library(DataComputing)
library(tidyverse)
```


```{r}
#Loading in all the necessary data

pk_stats <- read_csv(file = "pokemon_stats.csv")
pk_types <- read_csv(file = "pokemon_types.csv")
pk_names_extra <- read_csv(file = "pokemon.csv")
stat_names <- read_csv(file = "stat_names.csv")
type_names <- read_csv(file = "type_names.csv")
pk_stats
pk_types
pk_names
stat_names
type_names
```

```{r}
# Getting English type names
type_names <-
  type_names %>%
  filter(local_language_id == 9)
type_names
```

```{r}
# Getting English Stat Names
stat_names <-
  stat_names %>%
  filter(local_language_id == 9)
stat_names
```

```{r}
# Selecting the rows needed from the pokemon_names_extra var
pk_names <-
  pk_names_extra %>%
  select(species_id, identifier)
pk_names
```

```{r}
#Creating correct info

pk_stats_f <-
  pk_stats %>%
  left_join(y = pk_names, by = c("pokemon_id" = "species_id")) %>%
  select(identifier, stat_id, base_stat) %>%
  left_join(y = stat_names) %>%
  select(identifier, name, base_stat) %>%
  rename(stat = name, 
         name = identifier)
pk_stats_f
```

```{r}
pk_types_f <-
  pk_types %>%
  left_join(y=pk_names, by = c("pokemon_id" = "species_id")) %>%
  select(identifier, type_id, slot) %>%
  left_join(y = type_names) %>%
  select(identifier, name, slot) %>%
  rename(name = identifier, 
         type = name)
pk_types_f
```

```{r}
pk_stats_f %>%
  str()
```

```{r}
pk_types_f %>%
  str()
```

```{r}
# We can displaying the difference in stats for an individual pokemon

pk_stats_f %>%
  filter(name == "bulbasaur") %>%
  ggplot(aes(stat, base_stat)) + 
  geom_col() + 
  ggtitle("Bulbasaur's Stats") + 
  ylab("Stat Value") + 
  xlab("Stat Name")
```

```{r}
# We can filter pokemon by type
pk_types_f %>%
  filter(type == "Fire" & slot == 1)
```

```{r}
# Creating the master data
master <-
  pk_stats_f %>%
  left_join(pk_types_f)

# Example of Research Question: How do the average attack stats compare across types?

master %>%
  filter(stat == "Attack") %>%
  group_by(type) %>%
  summarise(avg_attack = mean(base_stat)) %>%
  ggplot(aes(x = type, y = exp(avg_attack))) + 
  geom_col() + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))

# As you can see from the chart, Fighting types have the best attack on average across types. 
```

## Sources

Our primary source is https://github.com/Akolyte/STAT-184-Final-Project/blob/master/pokemon.csv. There are other files that we may use that have different statistics, so we can join the datasets then if needed.
Another source that we are using is an API that has data regarding the different Pokemon games to see if we can make trends relate to specific games or times when they were released. This is found on www.pokeapi.co.
